Ionization of molecules by multicharged bared ions by using the stoichiometric model

### 2019-09-16 NewMonti25.tex
DNA figure in working progress

Then, by implementing the SSM and the CDW --instead of the first Born 
approximation--, we calculate ionization cross-section of several 
molecules (see Table~\ref{tab:families}) by the impact of antiprotons,
H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$,\ and O$^{+8}$. Furthermore, 
in Section 3, we investigate different DNA and RNA molecules
such as adenine, cytosine, guanine, thymine, uracil, and also DNA 
backbone.

@@

\section{Theory: Ionization of Atoms}

@@

In Figure 1, we report our total ionization results for the six 
essential elements by the impact of six different projectiles: 
antiprotons, H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
To reduce the resulting 36 magnitudes into a single consistent 
figure, we considered the fact that in the first Born approximation
the ionization cross-section scales with the square of the projectile 
charge, $Z^{2}$. The values of the impact energies considered 
range between 0.1 to 10 MeV/amu, where the CDW is supposed 
to hold. In fact, for the highest projectile charges the minimum 
impact energy where the CDW is expected to be valid could be 
higher than 100 KeV. Moreover, we corroborated that the first Born 
approximation provides quite reliable results only for energies higher 
than a couple of MeV/amu. We use the same color to indicate the 
projectile charge throughout the figures of this work: dashed-red, 
solid-red, blue, magenta, olive and orange for antiprotons, 
H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$, respectively. 
Notably, there is no complete tabulation of ionization of atoms by 
the impact of multicharged ions. We hope that the ones presented 
in this article will be of help for future works.
\begin{figure}
\includegraphics[width=\textwidth]{figuras/atomicscaling.eps}
\caption{.}
\label{eq:atomscaling}
\end{figure}

\section{Ionization of Molecules}

@@

Lets us consider a molecule $M$ composed by $n_{\alpha}$ atoms of the
element $\alpha$, the SSM describes the total ionization cross section 
of the molecule $\sigma_{M}$ as a simple sum of ionization cross 
sections of the isolated atoms $\sigma_{\alpha}$, 
\begin{equation}
 \sigma_{M}=\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}\,.  
 \label{eq:sumion}
\end{equation}
We divided sixteen molecular targets of our interest in three families. 
The classification defined is given in Table~\ref{tab:families}.
\begin{table}
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.62\textwidth}|}
\hline
 CH  & CH$_4$, C$_2$H$_2$, C$_2$H$_4$, C$_2$H$_6$, and C$_6$H$_6$ \\
\hline
 CHN & C$_5$H$_5$N, C$_4$H$_4$N$_2$, C$_2$H$_7$N, and CH$_5$N \\
\hline
 \multirow{4}{*}{DNA} & adenine (C$_5$H$_5$N$_5$), 
                        cytosine (C$_4$H$_5$N$_3$O), \\
     & guanine (C$_5$H$_5$N$_5$O), thymine (C$_5$H$_6$N$_2$O$_2$), \\
     & uracil (C$_4$H$_4$N$_2$O$_2$), tetrahydrofuran (C$_4$H$_8$O), \\
     & DNA backbone (C$_5$H$_{10}$O$_5$P), \\
     & dry DNA (C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$)\\
\hline
\end{tabular}
\caption{Seventeen molecular targets of our interest classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

In Figure 2, we report the total ionization cross sections by the 
impact of multicharged ions for the four fundamental DNA components: 
adenine, cytosine, guanine and thymine. Our reports for uracil and 
DNA backbone containing P are given in Figure 3. For adenine, the 
agreement with the experimental data available in our range of validity 
is quite good. For uracil, we have a puzzling situation: our results 
show good agreement with the experiments by proton impact of Itoh 
{\it et al.}~\cite{itoh2013}, but for the same target our theory fails 
by a factor of two for the impact of C$^{+4}$ and O$^{+6}$ ions. 
Nonetheless, it should be stated that our theoretical results coincide 
with calculations by Champion, Rivarola and 
collaborators~\cite{champion2012,agnihotri2012}, which may indicate a 
possible misstep of the experiments.

@@

\bibitem{itoh2013} 
A. Itoh, Y. Iriki, M. Imai, C. Champion, and R. D.
Rivarola, Cross sections for ionization of uracil by MeV-energy-proton
impact, 
Phys. Rev. A 88, 052711 (2013)

@@

>>>>

Then, by implementing the SSM and the CDW --instead of the first Born 
approximation--, we calculate ionization cross-section of several 
molecules (see Table~\ref{tab:families}) by the impact of antiprotons,
H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$,\ and O$^{+8}$. Furthermore, 
in Section 3, we investigate different DNA and RNA molecules
such as adenine, cytosine, guanine, thymine, uracil, tetrahydrofuran 
(THF), pyrimidine and also DNA backbone.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory: Ionization of Atoms}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[H]
\centering
%\includegraphics[width=\textwidth]{figuras/atomicscaling.eps}
\includegraphics[width=0.7\textwidth]{figuras/Fig_finales/Fig1.eps}
\caption{Scaled total ionization cross section of six atomic targets 
by impact of six projectiles.}
\label{eq:atomscaling}
\end{figure} 
In Figure 1, we report our total ionization results for the six 
essential elements by the impact of six different projectiles: 
antiprotons, H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
To reduce the resulting 36 magnitudes into a single consistent 
figure, we considered the fact that in the first Born approximation
the ionization cross-section scales with the square of the projectile 
charge, $Z^{2}$. The values of the impact energies considered 
range between 0.1 to 10 MeV/amu, where the CDW is supposed 
to hold. In fact, for the highest projectile charges the minimum 
impact energy where the CDW is expected to be valid could be 
higher than 100 KeV. Moreover, we corroborated that the first Born 
approximation provides quite reliable results only for energies higher 
than a couple of MeV/amu. We use the same color to indicate the 
projectile charge throughout the figures of this work: dashed-red, 
solid-red, blue, magenta, olive and orange for antiprotons, 
H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$, respectively. 
Notably, there is no complete tabulation of ionization of atoms by 
the impact of multicharged ions. We hope that the ones presented 
in this article will be of help for future works.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Ionization of Molecules}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Lets us consider a molecule $M$ composed by $n_{\alpha}$ atoms of the
element $\alpha$, the SSM describes the total ionization cross section 
of the molecule $\sigma_{M}$ as a simple sum of ionization cross 
sections of the isolated atoms $\sigma_{\alpha}$, 
\begin{equation}
 \sigma_{M}=\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}\,.  
 \label{eq:sumion}
\end{equation}
We divided sixteen molecular targets of our interest in three families. 
The classification defined is given in Table~\ref{tab:families}.
\begin{table}
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.56\textwidth}|}
\hline
 CH  & CH$_4$, C$_2$H$_2$, C$_2$H$_4$, C$_2$H$_6$, C$_6$H$_6$ \\
\hline
 CHN & C$_5$H$_5$N, C$_4$H$_4$N$_2$, C$_2$H$_7$N, CH$_5$N \\
\hline
 \multirow{5}{*}{DNA} 
     & C$_5$H$_5$N$_5$ (adenine), C$_4$H$_5$N$_3$O (cytosine), \\
     & C$_5$H$_5$N$_5$O (guanine), C$_5$H$_6$N$_2$O$_2$ (thymine), \\
     & C$_4$H$_4$N$_2$O$_2$ (uracil), C$_4$H$_8$O (THF), \\
     & C$_5$H$_{10}$O$_5$P (DNA backbone), \\
     & C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$ (dry DNA) \\
\hline
\end{tabular}
\caption{Seventeen molecular targets of our interest classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

In Figure 2, we report the total ionization cross sections by the 
impact of multicharged ions for adenine, cytosine, guanine and thymine. 
For adenine, the agreement with the experimental data 
available (circles)~\cite{iriki2011} is quite good in our range of validity. 
To the best of our knowledge, there are not experimental data on 
ion--collision ionization for the rest of the molecules. We included 
electron impact measurements (right--oriented triangles)~\cite{mozejko2003} 
for incident energies greater than 500keV, were good agreement is found 
with proton ionization for all the molecules. 
%In fact, recent models for electron impact in complex biomolecules 
%have been under progress~\ref{tan2018}.

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figuras/Fig_finales/Fig2.eps}
\caption{Present ionization cross section by impact of six projectiles, 
and experimental values for proton and electron ionization.}
\label{eq:crossDNA}
\end{figure}

Our reports for uracil, DNA backbone, pyrimidine and THF are given in 
Figure 3. For uracil, we have good agreement with the experimental 
proton impact measurements by Itoh~{\it et al.}~\cite{itoh2013}. 
However, for the same target, our theory fails by a factor of two for 
the experimental ionization values by the impact of C$^{+4}$~(square symbols), 
O$^{+6}$~(circle symbols) and O$^{+8}$~(triangle--down symbols) 
ions~\cite{agnihotri2012,agnihotri2013}. 
Nonetheless, it should be stated that our theoretical results coincide 
with calculations by Champion, Rivarola and 
collaborators~\cite{champion2012,agnihotri2012}, which may indicate a 
possible misstep of the experiments. 
For pyrimidine, we show comparison of our results with experimental 
proton~\cite{wolff2014} and electron~\cite{bug2017} ionization. 
The electron impact measurements agree with our calculations for energies
higher than 500keV. Unexpectedly, the proton impact cross sections are
significantly lower than our findings. 
The THF molecule results are compared with proton~\cite{wang2016} and 
electron~\cite{wolf2019,bug2017,fuss2009} impact measurements, showing
overall good agreement with our theory.

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{figuras/Fig_finales/Fig3.eps}
\caption{Present ionization cross section by impact of six projectiles, 
and experimental values for proton and electron ionization.}
\label{eq:crossDNA}
\end{figure}

@@

\bibitem{iriki2011}
Y. Iriki, Y. Kikuchi, M. Imai, and A. Itoh
Phys. Rev. A \textbf{84} 052719 (2011).

\bibitem{mozejko2003}
P. Mozejko and L. Sanche, Radiat Environ. 
Biophys \textbf{42}, 201 (2003).

\bibitem{tan2018}
H. Q. Tan, Z. Mi, and A. A. Bettiol, 
Simple and universal model for electron-impact ionization of complex biomolecules, 
Phys. Rev. E \text{97}, 032403 (2018)

\bibitem{itoh2013} 
A. Itoh, Y. Iriki, M. Imai, C. Champion, and R. D. Rivarola, 
Cross sections for ionization of uracil by MeV-energy-proton impact, 
Phys. Rev. A 88, 052711 (2013).

@@

\bibitem{agnihotri2013}
A N Agnihotri, S Kasthurirangan, S Nandi, A Kumar, C Champion,, H Lekadir, 
J Hanssen, P FWeck, M E Galassi, R D Rivarola, O Fojon and L C Tribedi, 
Absolute total ionization cross sections of uracil (C4H4N2O2) in 
collisions with MeV energy highly charged carbon, oxygen and fluorine ions
J. Phys. B \textbf{46}, 185201 (2013).

\bibitem{wolff2014}
W. Wolff, H. Luna, L. Sigaud, A. C. Tavares, and E. C. Montenegro
Absolute total and partial dissociative cross sections of pyrimidine
at electron and proton intermediate impact velocities
J. Chem. Phys. \textbf{140}, 064309 (2014).

\bibitem{bug2017}
M. U. Bug, W. Y. Baek, H. Rabus, C. Villagrasa, S. Meylan, A. B. Rosenfeld,
An electron-impact cross section data set (10 eV--1 keV) of DNA
constituents based on consistent experimental data: A requisite for 
Monte Carlo simulations,
Rad. Phys. Chem. \textbf{130} 459--479 (2017).

\bibitem{wang2016}
M. Wang, B. Rudek, D. Bennett, P. de Vera, M. Bug, T. Buhr, W. Y. Baek, 
G. Hilgers, H. Rabus, 
Cross sections for ionization of tetrahydrofuran by protons at energies 
between 300 and 3000 keV
Phys. Rev. A \textbf{93}, 052711 (2016).

\bibitem{wolf2019}
W. Wolff, B. Rudek, L. A. da Silva, G. Hilgers, E. C. Montenegro, 
M. G. P. Homem,
Absolute ionization and dissociation cross sections of tetrahydrofuran:
Fragmentation--ion production mechanisms
J. Chem. Phys. \textbf{151}, 064304 (2019).

\bibitem{fuss2009}
M. Fuss, A. Muñoz, J. C. Oller, F. Blanco, D. Almeida, P. Limão-Vieira, 
T. P. D. Do, M. J. Brunger, G. García,
Electron-scattering cross sections for collisions with tetrahydrofuran 
from 50 to 5000 eV
Phys. Rev. A \textbf{80}, 052709 (2009).
